---
title: "PA"
output: html_document
date: "2024-08-28"
---

Install necessary packages 
#Load the libraries
```{r}
install.packages(c("dplyr", "tidyr", "caret", "stringr", "mice"))
library(dplyr)
library(tidyr)
library(caret)
library(stringr)
library(mice)
library(ggplot2)
library(reshape2)
library(readxl)
```

#Load the dataset
```{r}
train_data <- read_excel("C:/Users/Admin/Downloads/train.xlsx")
View(train_data)
```


#---------------------------PREPROCESSING DATA---------------------------------


-///*HANDLING MISSING VALUES-------------------------------------------------------
#Check missing values in the entire dataset
```{r}
missing_values_summary <- sapply(train_data, function(x) sum(is.na(x)))
print(missing_values_summary)
```

# Impute missing values for numerical variables using median
Median for Numerical Variables
Mileage, Engine, Max Power, Seats:
Mileage (e.g., kmpl): Cars can have varying fuel efficiency, and using the median helps avoid skew from very high or very low values.
Engine (e.g., CC): Engine sizes can vary widely, and the median provides a typical value without being skewed by rare, extreme sizes.
Max Power (e.g., bhp): Similar to engine size, power ratings can have outliers; the median reflects a typical car's power.
Seats: Most cars have a standard number of seats; the median helps maintain a realistic average when data is missing.

```{r}
train_data$mileage <- ifelse(is.na(train_data$mileage), median(as.numeric(str_extract(train_data$mileage, "\\d+\\.\\d+")), na.rm = TRUE), train_data$mileage)
train_data$engine <- ifelse(is.na(train_data$engine), median(as.numeric(str_extract(train_data$engine, "\\d+")), na.rm = TRUE), train_data$engine)
train_data$max_power <- ifelse(is.na(train_data$max_power), median(as.numeric(str_extract(train_data$max_power, "\\d+")), na.rm = TRUE), train_data$max_power)
train_data$seats <- ifelse(is.na(train_data$seats), median(train_data$seats, na.rm = TRUE), train_data$seats)
View(train_data)
```


# Replace missing torque with the mode (most frequent value)
Standard Settings: Many cars share engine configurations, leading to repeated torque values. The mode captures this commonality.
Consistency: Using the mode ensures that missing torque values are filled with typical, realistic configurations seen in the dataset.
```{r}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
train_data$torque <- ifelse(is.na(train_data$torque), Mode(train_data$torque), train_data$torque)
View(train_data)
```



# CONVERTING NON-NUMERIC COLUMNS-----------------------------------------------
# Extract numerical part from mileage, engine, max_power
```{r}
train_data$mileage <- as.numeric(str_extract(train_data$mileage, "\\d+\\.\\d+|\\d+"))
train_data$engine <- as.numeric(str_extract(train_data$engine, "\\d+"))
train_data$max_power <- as.numeric(str_extract(train_data$max_power, "\\d+\\.\\d+|\\d+"))
train_data$torque <- as.numeric(str_extract(train_data$torque, "^\\d+"))
View(train_data)
```


# Convert categorical variables to factors
```{r}
train_data$fuel <- as.factor(train_data$fuel)
train_data$seller_type <- as.factor(train_data$seller_type)
train_data$transmission <- as.factor(train_data$transmission)
train_data$owner <- as.factor(train_data$owner)

# Convert categorical variables to factors with specified reference levels
train_data$fuel <- relevel(train_data$fuel, ref = "Diesel") 
train_data$seller_type <- relevel(train_data$seller_type, ref = "Individual")
train_data$transmission <- relevel(train_data$transmission, ref = "Manual")  
train_data$owner <- relevel(train_data$owner, ref = "First Owner") 

proportions(table(train_data$fuel))
proportions(table(train_data$seller_type))
proportions(table(train_data$transmission))
proportions(table(train_data$owner))

```


# FEATURE ENGINEERING----------------------------------------------------------
# Create a new column for the age of the car
```{r}
train_data$car_age <- 2024 - train_data$year
View(train_data)
```

# Remove columns that are irrelevant for prediction, such as 'name' and 'year'
# Do not run this code
```{r}
train_data <- train_data %>% select(-c(name, year))
```



#ENCODING CATEGORICAL VARIABLES------------------------------------------------
# Use one-hot encoding for categorical variables/ Create dummy variables for specific columns (phương án 1)
```{r}
dummy_model <- dummyVars(" ~ fuel + seller_type + transmission + owner", data = train_data)
train_data_dummies <- predict(dummy_model, newdata = train_data)
train_data_dummies <- as.data.frame(train_data_dummies)
```

#Use Label Encoding (phương án 2)
```{r}
train_data <- data.frame(
  fuel = factor(c("Petrol", "Diesel", "Petrol")),
  seller_type = factor(c("Dealer", "Individual", "Dealer")),
  transmission = factor(c("Manual", "Automatic", "Automatic")),
  owner = factor(c("First Owner", "Second Owner", "First Owner"))
)

# Convert categorical columns to numeric codes
train_data$fuel <- as.numeric(train_data$fuel)
train_data$seller_type <- as.numeric(train_data$seller_type)
train_data$transmission <- as.numeric(train_data$transmission)
train_data$owner <- as.numeric(train_data$owner)
```


# Remove original categorical columns from the original data
```{r}
train_data <- train_data %>%
  select(-c(fuel, seller_type, transmission, owner)) %>%
  bind_cols(train_data_dummies)
View(train_data)
```



# NORMALIZE/STANDARDIZE DATA---------------------------------------------------
# Standardize the numerical features
```{r}
summary(train_data)
pre_process <- preProcess(train_data, method = c("center", "scale"))
train_data <- predict(pre_process, train_data)
```


# FINAL CHECKS-----------------------------------------------------------------
# Check the structure of the processed data
```{r}
str(train_data)
```

# Check for any remaining missing values
```{r}
sum(is.na(train_data))

missing_values <- sapply(train_data, function(x) sum(is.na(x)))
missing_values[missing_values > 0]
missing_positions <- which(is.na(train_data$max_power))
```

There are 1 missing value in max_power col, row 4682

#Replace this missing value with the median(max_power)
```{r}
train_data$max_power <- ifelse(is.na(train_data$max_power),                               median(as.numeric(train_data$max_power), na.rm = TRUE),train_data$max_power)
sum(is.na(train_data))
```


# Save the Preprocessed Data
```{r}
write.csv(train_data, "C:/Users/Admin/Downloads/PA-Data_preprocessed.csv", row.names = FALSE)
```





#------------------------------MODEL BUILDING----------------------------------


